### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_mediation.tex
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1305 d8b0c59b-6544-0410-94d9-83a5658a058e

In D2.4.1 we reviewed current work and literature, discussing the main areas of the topic, that is the core problem of ontology mediation, the types of mismatches that occur between ontologies, and the three main ontology mediation approaches, namely (1) the representation of correspondences between ontologies, (2) the (semi-)automatic process of discovering these correspondences and (3) creation of a new ontology from the source ontologies.

Based on the theoretical foundations laid down in D2.4.1 we present our approach to integrating ontology matching into the FAST platform. We present the roles related to ontology matching in FAST, and the way ontology matching is achieved within each one of those roles.

@@

\begin{tabular}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
    \hline
    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {27/02/2011}\\ \hline
    \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
    \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.3 is the third iteration of the FAST ontology matching approach. The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for further analysis of the scenarios arising in FAST that require ontology mediation. Based on that general overview, this current iteration presents the approach to implement ontology matching within FAST. We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne,  \authorTwo, \authorThree, \authorFour, \authorFive}\\
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\
    \hline
    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {knud.moeller@deri.org} \\ \cline{4-5}
    \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
    \textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment}\\ \hline
\end{tabular}
\end{small}
\newpage

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{3}
\tableofcontents
\clearpage
\section*{Lists of Tables and Figures}
\listoftables
\listoffigures
\cleardoublepage
%\pagenumbering{arabic}

@@

We aim to integrate a form of ontology matching into the FAST platform to allow for discrepant systems to intercommunicate.  We try to find a solution to these differences which range from simple naming differences to more complex data heterogeneity, behavioural differences, discrepant paradigms and other disparities. The reconciliation of these differences is called ontology matching (often mentioned in the specialised literature as ontology mediation, or ontology alignment, thus we will sometimes use these terms as well).

@@

\subsection{Structure of the Document}
This document describes the approach of integrating ontology matching into the FAST environment. Sect.~\ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand. Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilities of the tool, and the testing procedure with the results. The remaining part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue. Sect.~\ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Sect.~\ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.

%Describe the rest.
The Appendix contains the first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching. We discuss ontologies in general in App.~\ref{sec:ontologies}, present the ontology mediation problem in App.~\ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in App.~\ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in App.~\ref{sec:mediationapproaches}.

@@

%\pagebreak
%\part{Implementing Ontology Matching within the FAST Platform}
%\label{part:d2.4.2}

@@

\section{Summary}
\label{sec:summary}
With the Semantic Web winning more and more ground both in research and industry, the mismatches between ontologies used by the inter-operating systems present an increasing problem. Ontology matching deals with identifying these mismatches, and overcoming them.

In the first iteration (described in the Appendix) we laid the theoretical foundations of ontology matching. After presenting the main problem, and the types of mismatches between ontologies, we have discussed - based on the most prominent research results in these areas - the three main approaches to ontology mediation, that is \textit{ontology mapping} (representing  correspondences between entities), \textit{ontology alignment} (the process of finding the correspondences between ontologies) and \textit{ontology merging }(the process of creating a new ontology based
on the source ontologies that need to be reconciled).

In this second iteration (described in the main body of this document) we established the approach to integrate ontology matching into FAST, and we have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level. Firstly, we have shown the implementation details as well as test scenarios, we have presented a tool for finding the alignments between ontologies as well as a powerful language to represent them, we have described a tool to compose mapping rules in a consistent manner with other FAST components and the means to save them in the Semantic Catalogue. Secondly, we described the way in which mapping operators are used in the GVS to combine discrepant screens. Lastly, we presented the way the matching operator and mapping rules are deployed and exploited in the JavaScript gadget.

For the last and final iteration of deliverable D2.4 we plan to evaluate real-life scenarios and finalise the implementation of ontology matching for the FAST platform.

>>>>

In \cite{ambrus2009mediation} we reviewed current work and literature, discussing the main areas of the topic: the core problem of ontology mediation, the types of mismatches that occur between ontologies, and the three main ontology mediation approaches, namely (1) the representation of correspondences between ontologies, (2) the (semi-)automatic process of discovering these correspondences and (3) creation of a new ontology from the source ontologies.

Based on these theoretical foundations, the second~\cite{ambrus2010mediation} as well as this final iteration of the deliverable present our approach to integrating ontology matching and the definition of mapping rules into the FAST platform. We present the roles related to ontology matching in FAST, and the way ontology matching is achieved within each one of those roles. We have developed and integrated a tool to allow end users to define mappings of varying complexity and create new operator building blocks to be used in the screen design process.

@@

\begin{tabular}
    {| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
    \hline
    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {27/02/2011}\\ \hline
    \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
    \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.3 is the third iteration of the FAST ontology matching and mapping approach. The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for further analysis of the scenarios arising in FAST that require ontology mediation. Based on that general overview, the second and third iteration presents the approach to implement ontology matching within FAST. We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role, including the \emph{fact transformation tool}, which has become part of the FAST tool chain.}\\ \hline
    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne,  \authorTwo, \authorThree, \authorFour, \authorFive}\\
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\ 
%   {} & \multicolumn{4}{l|}{}\\
    \hline
    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {knud.moeller@deri.org} \\ \cline{4-5}
    \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
    \textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment}\\ \hline
\end{tabular}
\end{small}
\newpage

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{3}
\tableofcontents
\clearpage
% \section*{Lists of Tables and Figures}
\listoftables
\listoffigures
\cleardoublepage
%\pagenumbering{arabic}

@@

We aim to integrate a form of ontology matching into the FAST platform to allow for discrepant systems to intercommunicate --- or, more precisely, previously incompatible building blocks in our case.  We try to find a solution to these differences which range from simple naming differences to more complex data heterogeneity, behavioural differences, discrepant paradigms and other disparities. The reconciliation of these differences is called ontology matching (often mentioned in the specialised literature as ontology mediation, or ontology alignment, thus we will sometimes use these terms as well).

While we have explored ways of automatically defining mappings between different conceptualisations, our focus within FAST is less on research into new such methods. Instead, we aim to support the end user in defining such mapping rules within the FAST tool chain and integrate the process within the platform. To this end, we have developed and integrated a fact transformation tool, which is part of the FAST GVS and presented in this deliverable.

@@

\subsection{Structure of the Document}
This document describes the approach of integrating ontology matching into the FAST environment. Sect.~\ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand. Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilities of the tool, and the testing procedure with the results. The remaining part of the section deals with the \emph{fact transformation tool}, developed to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue. Sect.~\ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Sect.~\ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.

%Describe the rest.
The Appendix contains the main content of first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching. We discuss ontologies in general in App.~\ref{sec:ontologies}, present the ontology mediation problem in App.~\ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in App.~\ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in App.~\ref{sec:mediationapproaches}.

@@

@@

### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part1.tex
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1305 d8b0c59b-6544-0410-94d9-83a5658a058e

@@

The first row (reference) shows the reference alignment, which, naturally, has both perfect precision and recall. We can observe what intuition has predicted, namely that pure string equality (equality) is far too simple and irrelevant, by only taking identical labels. By using string distances and giving certain thresholds (Levenshtein and SMOA), we can see that the results are much less precise, but have a better recall, since this allows for entities having similar names to be discovered, at the expense of having quite a few incorrect results (lower precision); the thresholds allow for low-scored cases to be eliminated, although this results in the exclusion of some correct correspondences. The last column (JWNL) contains the results of the Wordnet-enabled method, which shows quite an improvement (precision of 0.67 and a recall of 0.75\footnote{The scores are orientative, giving a general picture of how the process approaches 100\% efficiency}), due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results. The precision of the JWNL alignment shows only a tiny drop below the recall value, meaning that the number of incorrect correspondences discovered is small, and the main source of error is from the number of correspondences not discovered.

@@

To support manual ontology mapping, we have developed a rule based, interactive Facts Transformation Tool. This tool utilizes techniques of our service wrapper tool, cf. deliverable D4.3.2. One part of the service wrapper tool deals with a rule based approach for the transformation of server responses into FAST facts. This problem is closely related to the problem of ontology mediation. Thus, these techniques also allow to create data transformation operators, i.e. operators that mediate between two ontologies.

Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a screen shot of the  Facts Transformation Tool within the GVS environment. The Facts Transformation Tool has three tabs. The \textit{General} tab allows to define the pre- and postconditions for the desired data transformation operator. These conditions are  typed according to the participating semantic elements to be derived from the FAST catalogue. This typing will provide the structure of the input and output data of the desired data transformation operator.

Figure \ref{fig:FactsTransformationToolTransformationRules} shows the editor for transformation rules. On the left of Figure \ref{fig:FactsTransformationToolTransformationRules} there is a tree of example data that is used to validate the transformation rules during editing. Thus, the Facts Transformation Tool will execute the transformation rules after any editing in order to give direct feedback on their functionality and to enable the user to validate the achieved transformation. The transformation results are shown on the right of Figure \ref{fig:FactsTransformationToolTransformationRules}.

@@

The middle of Figure \ref{fig:FactsTransformationToolTransformationRules} shows some possible transformation rules for our example. A \textit{createObject} rule may be used to create output fact objects. The \textit{Target} field of such a rule specifies the type of the created objects, in our example \textit{List}. The top rule fires for each element of the input data with the type specified in the \textit{From} field of the transformation rule. Thus, in our example, the input data is searched for objects of type \textit{List} firing for the root of our data.

The second rule of our example creates \textit{Customer} objects. This rule is a subrule of the root rule and thus the created objects are added to the \textit{List} object created by the parent rule. Similarly, the search for triggering objects of type \textit{Person} is restricted to the subtree below the from element that has triggered the parent rule, i.e. to our input \textit{List}.

The third rule is used to fill the attribute \textit{fullName} of the \textit{Customer} objects. This rule is an example for a many to one transformation. The result is a string concatenation of the \textit{lastName} of the \textit{Person} that has triggered the parent rule plus a text constant plus the \textit{firstName} of the current \textit{Person}. The resulting string is transferred to the \textit{fullName} attribute of the corresponding \textit{Customer}.

Note that simple one-to-one transformation rules may be derived with the help of the ontology mediation approach described in Section \ref{subsubsec:automaticmapping}. Such simple rules may be added to the transformation tool, automatically. We try to add this functionality within the remaining project time. The more complex rules may then be added interactively. During editing, the semantics provided by the FAST catalogue is used to support the user with completion proposals and consistency checking. In addition, the direct execution of the rules on example data helps the user to create appropriate rules and to validate their effects.

@@

Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST GVS tool. For this purpose a simple semantic description is sent to the FAST catalogue defining the semantics of the operator and the semantics of its pre- and postconditions. This is shown in Fig.~\ref{fig:operator_in_screen_builder}, where the newly created operator for mediating between Person and Customer concepts (i.e., transforming instances of one concepts into instances of the other) has been saved and is now available in the GVS screen builder interface.

@@

The Service Wrapper Tool uses its own repository for trees of transformation rules. First of all this has technical reasons: the development of the transformation tool is based on a reference architecture and libraries for web based interactive systems provided by Kassel University, cf. \cite{AsDH2009CVSM}. This repository has the special advantage that it provides support for teams of developers, thus the service wrapper tool allows to merge the contributions of multiple developers.

\section{Gadget Building Phase}
\label{sec:gadgetbuilding}
Every time ontology matching is performed and an alignment is created between two underlying resources, the corresponding alignment rules are stored and an operator is created in the catalogue.

At the gadget building phase, whenever preconditions and postconditions of screens, resources or operators show a semantic mismatch, the catalogue offers the possible matching operators that would solve the incompatibility problem. This way a matching operator can be placed between two pre- postconditions that would otherwise be incompatible, making interoperation between the two possible. In seldom cases, the catalogue may fail to retrieve an appropriate mediation operator. In such cases, the possibility of building a new matching operator with the help of the Facts Transformation Tool is presented, in order to meet the user's need.

\section{Runtime} 
\label{sec:runtime}
For deployment, the gadgets implementation is generated in Javascript. In addition, a small FAST runtime library (again Javascript) is added to the gadget. This runtime library provides notification mechanisms and manages the fact base of the gadget. For data transformation we generate Javascript code that connects itself to the fact base and any time a fact is sent to the precondition of the operator, it does the data transformation and sends the result to the postcondition. The code generated from the data transformation rules follows a recursive pattern where each rule becomes a Javascript function with parameters for the source and target fact subtrees to be transformed. The function body then uses code to search the source tree for sub elements that match the \textit{From} part of the rule and then code for the desired operation is generated. For debugging purposes the user may examine the code generated for an operator in the \textit{CodeGen Viewer} tab of the Facts Transformation tool.

Our code generation for service wrappers and transformation operators uses a template based approach. Thus, the \textit{CodeGen Viewer} tab will also allow to examine the templates used for code generation and to adapt the templates in order to meet changing APIs or to meet new requirements. Thereby, the code generation is not hard wired into our tool but may be adapted by (skilled) end users on the fly.

>>>>

In this section, we present a number of tools that can be used to help the ontology engineer in creating mapping rules to mediate between different conceptualisations. Some ontologies relevant for an e-commerce use case are introduced, and their automatic mapping is evaluated using one of the tools. The evaluation shows that automatic mapping is feasible only in certain cases, while other cases require a manual approach. The section concludes with the presentation of a tool developed in FAST to facilitate the use of (either automatically or manually created) mapping rules within the FAST tool chain.

@@

The first row (reference) shows the reference alignment, which, naturally, has both perfect precision and recall. We can observe what intuition has predicted, namely that pure string equality on term labels is far too simple and therefore irrelevant. By using string distances and giving certain thresholds (``Levenshtein and SMOA''), we can see that the results are much less precise, but have a better recall, since this allows for entities having similar names to be discovered, at the expense of having quite a few incorrect results; the thresholds allow for low-scored cases to be eliminated, although this results in the exclusion of some correct correspondences. The last row (``JWNL'') contains the results of the Wordnet-enabled method, which shows quite an improvement (precision of 0.67 and a recall of 0.75),
% \footnote{The scores are orientative, giving a general picture of how the process approaches 100\% efficiency}
due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results. The precision of the JWNL alignment shows only a tiny drop below the recall value, meaning that the number of incorrect correspondences discovered is small, and the main source of error is from the number of correspondences not discovered.

@@

To support the application of previously created mapping rules, we have developed a rule based, interactive \emph{fact transformation tool} (or \emph{fact tool} for short). This tool utilises techniques of our service wrapper tool. One part of the service wrapper tool~\cite{rivera2011connecting} uses a rule-based approach for the transformation of web service responses into FAST facts (i.e., instances of ontology concepts). The application of ontology mediation rules can be modelled in a similar vein. In the fact tool, we therefore use analogous rule-based techniques to create data transformation operators, i.e., operators that mediate between two different conceptualisation.

@@

Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a screen shot of the  fact transformation tool within the GVS environment. The tool has four tabs\footnote{The \texttt{CodeGen Viewer} tab is explained later in Sect.~\ref{sec:runtime}.}, of which the \textit{General} tab (currently shown in the screen shot) is employed to define the pre- and postconditions for the desired data transformation operator in a first general step. 
Using the terminology introduced previously, this allows to define a basic \emph{level 0} mapping between two concepts (\texttt{PersonList} and \texttt{CustomerList} in the example).
The fact tool communicates with the FAST catalogue to select concepts and provide example instances (as JSON data structures) that help the user to perform the correct choice.
% These conditions are typed according to the participating semantic elements to be derived from the FAST catalogue. This typing will provide the structure of the input and output data of the desired data transformation operator.  
After this step, the mapping can be refined to a higher level by defining concrete transformation rules. These can be edited in the \emph{exTransformation} tab as shown in Fig.~\ref{fig:FactsTransformationToolTransformationRules}. On the left of the tab the user is presented with a tree of example data for the input concept, while the right side of the tab shows a similar tree with the derived data for the output concept. This way, the user can continuously observe the effect of changes in the rule they are currently editing. The fact tool constantly executes the transformation rules after any editing, giving direct feedback on their effect and thus enabling the user to validate the achieved transformation.

Examples for the rules themselves are shown in the rule tree in the middle of Fig.~\ref{fig:FactsTransformationToolTransformationRules}. Starting with a root rule, additional rules can be added as child-rules, while existing rules can be moved around the tree or deleted. Rules can have different types, such as \emph{createObject} or \emph{fillAttributes}. A \textit{createObject} rule can be used to create output fact objects. The \textit{Target} field of such a rule specifies the type of the created objects (i.e., a concept from the FAST catalogue), in our example \textit{List}. The top rule fires for each element of the input data with the type specified in the \textit{From} field of the transformation rule. Thus, in our example, the input data is searched for objects of type \textit{List} firing for the root of our data. The second rule of our example creates \textit{Customer} objects. This rule is a sub-rule of the root rule and thus the created objects are added to the \textit{List} object created by the parent rule. Similarly, the search for triggering objects of type \textit{Person} is restricted to the subtree below the from element that has triggered the parent rule, i.e. to our input \textit{List}. The third rule is used to fill the attribute \textit{fullName} of the \textit{Customer} objects. This rule is an example for a many-to-one transformation. The result is a string concatenation of the \textit{lastName} of the \textit{Person} that has triggered the parent rule plus a text constant plus the \textit{firstName} of the current \textit{Person}. The resulting string is transferred to the \textit{fullName} attribute of the corresponding \textit{Customer}.

Note that simple one-to-one transformation rules may be derived with the help of the ontology mediation approach described in Sect.~\ref{subsubsec:automaticmapping}. Such simple rules can be added to the transformation tool, automatically. The more complex rules may then be added interactively. During editing, the semantics provided by the FAST catalogue are used to support the user with completion proposals and consistency checking. In addition, the direct execution of the rules on example data helps the user to create appropriate rules and to validate their effects.

@@

Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST GVS tool. For this purpose a simple semantic description is sent to the FAST catalogue defining the semantics of the operator and the semantics of its pre- and postconditions. This is shown in Fig.~\ref{fig:operator_in_screen_builder}, where the newly created data transformator for mediating between Person and Customer concepts (i.e., transforming instances of one concepts into instances of the other) has been saved and is now available in the GVS screen builder interface as an filter operator.

@@

The Service Wrapper Tool uses its own repository for trees of transformation rules. First of all this has technical reasons: the development of the transformation tool is based on a reference architecture and libraries for web-based interactive systems provided by Kassel University~\cite{AsDH2009CVSM}. This repository has the special advantage that it provides support for teams of developers, thus the service wrapper tool allows to merge the contributions of multiple developers.

\section{Gadget Building  and Runtime}

Neither gadget (or screenflow) building or runtime are directly involved with the ontology mapping process. However, the steps performed previously affect these phases of the lifecycle as well, as discussed briefly in the sections below.

\subsection{Gadget Building} % (fold)
\label{sec:gadgetbuilding}

Every time ontology matching is performed and an alignment is created between two underlying resources, the corresponding alignment rules are stored and an operator building block is created in the catalogue, to be available during screen development.

During the gadget building phase, whenever pre- and post-conditions of screens, resources or operators show a semantic mismatch, the catalogue offers the possible matching operators that would solve the incompatibility problem. This way a matching operator can be placed between two pre- and postconditions that would otherwise be incompatible, making interoperation between the two possible. In seldom cases, the catalogue may fail to retrieve an appropriate mediation operator. In such cases, the possibility of building a new mapping operator with the help of the fact transformation tool is presented, allowing the user to assume the role of ontology engineer in special cases.

\subsection{Runtime} 
\label{sec:runtime}
For deployment, the gadget's implementation is generated as an executable bundle containing a portion of Javascript code. In addition, a small FAST runtime library (again Javascript) is added to the gadget. This runtime library provides notification mechanisms and manages the fact base of the gadget. For data transformation purposes (i.e., ontology mapping) we generate Javascript code that connects itself to the fact base. Any time a fact is sent to the pre-condition of the operator, it performs the data transformation and sends the result to the post-condition. The code generated from the transformation rules follows a recursive pattern, where each rule is represented as a Javascript function with parameters for the source and target fact subtrees to be transformed. The function body then uses code to search the source tree for sub elements that match the \textit{From} part of the rule and then code for the desired operation is generated. For debugging purposes the user can examine the code generated for an operator in the \textit{CodeGen Viewer} tab of the fact tool.

Our code generation for service wrappers and transformation operators uses a template-based approach. Thus, the \textit{CodeGen Viewer} tab will also allow to examine the templates used for code generation and to adapt the templates in order to meet changing APIs or to meet new requirements. Thereby, the code generation is not hard wired into our tool but may be adapted by (skilled) end users on the fly.

\section{Summary}
\label{sec:summary}
With the Semantic Web winning more and more ground both in research and industry, the mismatches between ontologies used by the inter-operating systems present an increasing problem. Ontology matching deals with identifying these mismatches, and overcoming them.

The initial work on the ontology matching comprised an overview of the area in general, and was reflected in the first iteration of this deliverable (\cite{ambrus2009mediation}, now in App.~\ref{sec:background}). In it, we discussed the theoretical foundations of ontology matching. After presenting the main problem and the types of mismatches between ontologies, we presented --- based on a study of the state-of-the-art in these areas --- the three main approaches to ontology mediation, i.e., \textit{ontology mapping} (representing  correspondences between entities), \textit{ontology alignment} (the process of finding the correspondences between ontologies) and \textit{ontology merging} (the process of creating a new ontology based on the source ontologies that need to be reconciled).

The second iteration of the document~\cite{ambrus2010mediation} saw a positioning of ontology matching within the FAST project, both in terms of the FAST lifecycle and its associated user roles. It was defined which phases are affected in which way, and which actions by particular user roles are involved. We have shown how matching is done at the different phases of the gadget life cycle and the approach taken at every level. Additionally, we covered different candidates for tools to support automatic matching algorithms and evaluated the most promising ones on the backdrop of an e-commerce scenario. It was found that, while automatic matching is feasible for some, mostly simple cases, the manual definition of mapping rules will still be necessary for many real-life cases. Research into new methods for automatic matching of complex cases was deemed outside the scope of this project. Instead, the focus was on supporting users of the FAST tool to manually define mapping rules in an intuitive fashion. To this end, a first standalone prototype of the \emph{fact transformation tool} was developed and presented briefly.

This final iteration of the deliverable focussed mainly on the improvement of the fact tool and its transformation of a standalone tool into an integral part of the FAST tool chain, thereby supporting users and allowing them to use the same kind of interaction and user interface throughout the platform. The fact tool presented in this deliverable and the service wrapping tool presented in \cite{rivera2011connecting} therefore use a similar graphical approach. Without leaving the FAST platform, users can now assume the role of ontology engineer and define mapping rules of varying complexity, which can then be saved as operator building blocks, to be used in the screen development process. To allow this, the fact tool communicates with the FAST catalogue to guide users in the rule creation process and outputs definitions according to the FAST gadget ontology, as well as code to be used in the deployed gadget.
